Does Anybody Really Know What Time It Is?\\ Automating the Extraction of Date Scalars

### 2015-07-06 intro.tex
Added first two figures

\subsubsection{	Scalar Dates}
The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}. They are typically implemented as fixed-point types containing an offset from some epoch (\textit{e.g.} Julian Days.) This makes them compact to store and allows some temporal operations to be implemented very efficiently using arithmetic. Thus from the RDBMS perspective, representing dates in scalar form provides numerous benefits for users, both in terms of available operations and query performance.

@@

Categorical dates have a natural hierarchy associated with them generated by calendar binning. The visualization in Figure I1 shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.

>>>>

\subsubsection{	Scalar Dates}

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{figures/FigureI1}
\caption{Categorical Date Scalars.}
\label{fig:I1}
\end{figure}

The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}. They are typically implemented as fixed-point types containing an offset from some epoch (\textit{e.g.} Julian Days.) This makes them compact to store and allows some temporal operations to be implemented very efficiently using arithmetic. Thus from the RDBMS perspective, representing dates in scalar form provides numerous benefits for users, both in terms of available operations and query performance.

@@

Categorical dates have a natural hierarchy associated with them generated by calendar binning. The visualization in Figure \ref{fig:I1} shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{figures/FigureI2}
\caption{Quantitative Date Scalars.}
\label{fig:I2}
\end{figure}

Quantitative dates are typically used for time series on an axis that maps the underlying distance measure to display pixel distance. The quantitative analog to categorical binning is truncation whereby all the dates in a bin are represented by the first date in the bin. Truncation accurately preserves the distance semantics while enabling roll-up to coarser levels of detail. The visualization in Figure \ref{fig:I2} shows the same sales data in a time series rolled up to the quarter level.

These types of visualizations are difficult to specify with simple categorical string representations of dates. Thus the conversion of unparsed strings to date scalars is an important component of data modeling for analytics.
